---
title: "Problem Set 6: Sampling from a Voter File "
author: Your Name Here
output: pdf_document
---


```{r setup, include = FALSE}
options(width = 100)
knitr::opts_chunk$set(error = TRUE)
```

## Background

In this exercise, we will focus on sampling and sampling distributions when we have access to an entire census for a given population. In this case, the `data/durham.csv` file contains anonymized data on all registered voters in Durham County, NC that were registered to vote on election day of the 2020 presidential election. The variables in this dataset are:

| Name               | Description                                                                                                                    |
|--------------------|--------------------------------------------------------------------------------------------------------------------------------|
| `ncid`         | unique voter identification number                                                                                        |
| `reg_year`         | did person vote (1) or not (0) in 1994?                                                                                        |
| `age`              | age of registered voter                                                                                                        |
| `gender`           | voter gender identity (`"F"` = Female, `"M"` = Male, `"U"` = Undeclared)                                                       |
| `race`             | voter racial identity  (see table below)                                                                                       |
| `latino`           | voter identifies as Hispanic or Latino (`"HL"`), does not (`"NL"`), or undesignated (`"UN"`)                                   |
| `party`            | party registration (`"DEM"`  = Democrat, `"Rep"` = Republican, `"LIB"` = Libertarian, `"GRE"` = Green, `"UNA"` = Unaffiliated) |
| `drivers_lic`      | voter has drivers license (1) or not (0)?                                                                                      |


North Carolina provides the voter history data in a separate file with a row for each voter who actually casts a ballot in a given election. In the `data/vote_history.csv` file is a row for every voter that cast a ballot in the 2020 general election. The variables in this dataset are:

| Name               | Description                               |
|--------------------|-------------------------------------------|
| `ncid`             | unique voter identification number        |
| `vote_method_2020` | method of voting in 2020 general election |
 

For the purposed of this exercise, we will treat the voter list from `durhman.csv` as the population of interest. Doing so is an increasingly common approach for polling, where pollsters are now using the voter file as a sampling frame to conduct their polls. We will repeated sample from this population to better understand sampling uncertainty. 


**Note:** please follow the directions carefully about setting the seed for the sampling based questions. 


In the Durham data, the racial category codes are:

| Race | Description                           |
|------|---------------------------------------|
| A    | ASIAN                                 |
| B    | BLACK or AFRICAN AMERICAN             |
| I    | AMERICAN INDIAN or ALASKA NATIVE      |
| M    | TWO or MORE RACES                     |
| O    | OTHER                                 |
| P    | NATIVE HAWAIIAN or PACIFIC ISLANDER   |
| U    | UNDESIGNATED                          |
| W    | WHITE                                 |





\newpage

\newpage

## Question 1 (6 points)

Load the voter list data (`data/durham.csv`) into R using `read_csv` and save the data as `durham` and use the same function to load in the voter history data (`data/vote_history.csv`) as `vote_history`. Use `left_join` to merge the voter history data into the voter list data so that all rows of the voter list remain. Use the `ncid` variable to join the data sets and save this merged data as `durham`. 

Any registered voter who did not turn out in the 2020 general election will not appear in the `vote_history` data so their `vote_method_2020` variable will be missing in the merged data. Create a new variable called `vote_gen_2020` that is 1 if the voter turned out and 0 otherwise (the `is.na()` function may be helpful). 

In the write-up, state how many units are in the population (that is, how many rows are in the `durham` data) and the proportion of registered voters that actually voted in the 2020 elections. 

 **Rubric:** 1pt for Rmd file compiling (autograder); 3pt for data loaded and merged (autograder), 1pt for creation of the `vote_gen_2020` variable (autograder); 1pt for number of rows and turnout proportion reported (PDF). 


## Answer 1

```{r}
library(tidyverse)
durham <- read_csv('data/durham.csv', show_col_types = FALSE)
vote_history <- read_csv('data/vote_history.csv')
durham <- left_join(durham, vote_history, by = "ncid")
durham <- durham |>
  mutate(vote_gen_2020 = ifelse(is.na(vote_method_2020),0,1))
num_units <- nrow(durham)
prop_voted <- mean(durham$vote_gen_2020)
num_units
prop_voted
```
There are 231747 rows, meaning 231747 units in the population. The turnout proportion is 0.7228702. 

\newpage

## Question 2 (4 points)

Create a density histogram of age with a bin width of 1 and save this plot as `age_hist` (use the aesthetic mapping `y = after_stat(density)` in to accomplish this). Create a barplot  for turnout (`vote_gen_2020`) with the proportion on the y-axis (use the aesthetic mapping `y = after_stat(prop)` in `geom_barplot` to achieve this). For a slightly nicer plot, use the `width = 1` argument. Make sure both of these plots are shown in the PDF. 


 **Rubric:** 2pts for `age_hist` (autograder); 2pts for `turnout_hist` (autograder). 

## Answer 2
```{r}
library(ggplot2)
age_hist <- ggplot(durham, aes(x = age)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 1)
age_hist
turnout_hist <- ggplot(dat = durham, mapping = aes(x = vote_gen_2020)) +
  geom_bar(aes(y = after_stat(prop)), binwidth = 1)
turnout_hist
```


\newpage

## Question 3 (5 points)

Use `summarize()` to calculate the population mean and standard deviation of `age` and `vote_gen_2020` (that is the mean and SD of these variables in the `durham` data) and save the resulting tibble as `pop_parameters` with the tibble output looking like:

```
# A tibble: 1 × 4
  age_pop_mean age_pop_sd turnout_pop_prop turnout_pop_sd
         <dbl>      <dbl>            <dbl>          <dbl>
1         XX.X       XX.X            X.XXX          X.XXX
```

Make sure that the column names are the same for the autograder. (**Hint:** you can summarize multiple variables in the same call to `summarize`.) Use `knitr::kable()` to present the values in nicely formatted table with `digits = 2` to create nicely rounded numbers and informative column names (they may need to be abbreviated to fit on the page). 

 **Rubric:** 4pts for correct `pop_parameters` tibble (autograder); 1pts for nicely formatted table (PDF). 

## Answer 3
```{r}
library(knitr)
pop_parameters <- durham |>
  summarize(
    age_pop_mean = mean(age, na.rm = TRUE),
    age_pop_sd = sd(age, na.r = TRUE),
    turnout_pop_prop = mean(vote_gen_2020, na.rm = TRUE),
    turnout_pop_sd = sd(vote_gen_2020, na.rm = TRUE)
    )
kable(pop_parameters, digits = 2, caption = "pop. parameters for age and turnout")
```



\newpage

## Question 4 (8 points)


In the first line of the code chunk for this question use the following code:

```{r eval=FALSE}
library(infer)
set.seed(02138)
```

Then use `rep_slice_sample()` to take 1,000 samples of size 200 from `durham` and calculate the sample mean of `age` and the sample mean/proportion of turnout for each of these samples. Save these as variables named `age_mean` and `turnout_prop` and save the resulting tibble as `samples_n200`.  


Create a density histogram of the age means and save this as `age_mean_hist` (use a bin width of 0.5). Create a density histogram of the turnout proportions and save this as `turnout_prop_hist` (use a bin width of 0.005) (**NOTE: this is not a bar plot like in Question 2**). Make sure both of these plots are shown in the PDF and that they have informative labels. 

In the write-up, compare the sampling distribution of the sample mean and sample proportion here to the population distributions from question 2. Are the shapes of the sampling distributions similar or different to the population distributions? If different, how are they different? 


 **Rubric:** 2pts for correct `sample_n200` output (autograder); 2pts for correct `age_mean_hist` (autograder); 2pts for correct `turnout_prop_hist` (autograder); 1pt informative labels (PDF); 1pt comparison to population distributions (PDF). 

## Answer 4

```{r}
library(infer)
set.seed(02138)
samples_n200 <- durham |>
  rep_slice_sample(n = 200, reps = 1000) |>
  group_by(replicate) |>
  summarize(age_mean = mean(age),
            turnout_prop = mean(vote_gen_2020))
age_mean_hist <- ggplot(samples_n200, mapping = aes(x = age_mean)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.5)
age_mean_hist
turnout_prop_hist <- ggplot(samples_n200, mapping = aes(x = turnout_prop)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.005)
turnout_prop_hist
```

There are both similarities and differences between histograms fro q2 and from this question. As we can see in this histogram, we have an evenly distributed histogram at around 46 years old and it is skewed right. However, the age mean is almost the same at 46. Meanwhile, the turnout prop histogram shows a concentration of turnout between .7 and .75 . The biggest difference however, sees to be in the standart deviation, with it being around 19 for the mean age. The sampling distributions of means are around the same value as the population mean, but standard deviation changes due to sampling variability. 

\newpage

## Question 5 (7 points)


Use the `summarize()` function on `samples_n200` to calculate the average (named `ev_age` and `ev_turnout`) and standard deviation (named `se_age` and `se_turnout`) of each sample mean/proportion across the repeated samples. Save this tibble as `samp_dist_summary` and it should look like this: 
```
# A tibble: 1 × 4
  ev_age se_age ev_turnout se_turnout
   <dbl>  <dbl>      <dbl>      <dbl>
1   X.XX   X.XX      X.XXX      X.XXX
```

Make sure that the column names are the same for the autograder. Use `knitr::kable()` to present the values in nicely formatted table with `digits = 2` to create nicely rounded numbers. 

Compare the mean and SD of these sampling distributions to the population means and SDs from the previous question. Are these distributions centered on the same value? Which has more spread, the population distribution of age/turnout or the sampling distributions of their means? 


 **Rubric**: 4pts for correct `samp_dist_summary` tibble (autograder); 1pt for nicely formatted table (PDF); 2pts for discussion (PDF). 

## Answer 5
```{r}
library(purrr)
samp_dist_summary <- samples_n200 |>
  summarize(ev_age = mean(age_mean),
            se_age = sd(age_mean),
            ev_turnout = mean(turnout_prop), 
            se_turnout = sd(turnout_prop)
  )
kable(samp_dist_summary, digits = 2)
```

The most significant contrast lies in the dataset's variability. For instance, the data in this graph illustrates a standard deviation of 1.34 and an average age of 45.99. The standard deviation for the proportion of individuals turning out to vote is 0.03, with an average turnout proportion of 0.72. These statistics align with my visual assessment of the previous graphs. As previously emphasized, the standard deviation for age in the entire population is 19.0, while the average mean age of the total population is 46. The standard deviation in the original voter turnout proportion is 0.45, and the average turnout proportion is 0.72. The means were consistent, but the standard deviations were not.




\newpage

## Question 6 (5 points)

Now suppose we received a bad voter file that only contained voters under the age of 30 at the time of the 2020 election. Set the seed using `set.seed(02138)` and repeat the replicate sampling from Question 3 to take 1,000 samples of size 200 from `durham` after removing voters 30 and over. Calculate the sample proportion of turnout for each of these samples. Save this as a variable named `turnout_prop` and save the resulting tibble as `young_samples_n200`.  


Create a density histogram of the turnout proportions with informative labels and add a vertical line at the true population average turnout (of all ages). Save this plot as `young_turnout_prop_hist`.

Does the sampling distribution of turnout in these samples appear biased or unbiased for the population average turnout? Explain why based on the sampling process. 


  **Rubric:** 2pts for correct `young_samples_n200` tibble (autograder); 1pt for correctly identifying biased/unbiased (PDF); 2pt explaining bias/unbiasedness (PDF). 

## Answer 6
```{r}
durham_under_30 <- durham |>
  filter(age<30)
set.seed(02138)
young_samples_n200 <- replicate(1000, {
  sample_n(durham_under_30, size = 200) |>
    summarize(turnout = mean(vote_gen_2020))
}, simplify = FALSE) |>
  bind_rows() |>
  rename(turnout_prop = turnout)
true_avg_turnout <- mean(durham$vote_gen_2020)
young_turnout_prop_hist <- ggplot(young_samples_n200, aes(x = turnout_prop)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.01) +
  geom_vline(aes(xintercept = true_avg_turnout), color = "pink", linetype = "dashed", linewidth = 2) +
  labs(x = "turnout prop.", y = "density") + theme_minimal()
young_turnout_prop_hist
```
Bias is present in this new distribution, primarily due to two key factors.If researchers were to solely rely on this sample distribution to model the overall population, it would result in inaccuracies due to the following reasons because this sample exhibits an estimated difference of approximately 0.10 in turnout proportion compared to the broader population, and because the variation within this sample does not accurately represent the voting behaviors of the entire population. This underscores the fact that relying solely on the proportions of voters under 30 would yield a biased perspective of the voting patterns within the overall population.


\newpage



# Code Printout (DO NOT EDIT BELOW THIS)

```{r print-code, ref.label=knitr::all_labels(), echo = TRUE, eval=FALSE}

```
